Report: the non-equilibrium molecular dynamics simulation

### 2013-03-26 writedown/npt-neq/alanine.tex
improve

\section{Alanine dipeptide
  under turning on electric field}
\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig.ala/ele.field/field.eps}
  \caption{The maganitude of the external electric field as
    a function of time.}
  \label{fig:tmp3}
\end{figure}

We study the response of the configuration of the alanine dipeptide
with respect to a turning on electric field. The direction of this
field is along the $x$ dirction and the magnitude with respect to time
is plotted in Fig.~\ref{fig:tmp3}. From $t=0$ to
$t=t_{\textrm{init}}$, the strength of the field grows linearly, while
after $t=t_{\textrm{init}}$, the strength is kept constant at
$E_{\textrm{max}}$.  The size of the system is $2.67\times 2.67\times
2.67\, \textsf{nm}^3$, with one alanine dipeptide solved in 644 TIP3P
water molecules. Firstly, an equilibrium NVT simulation of
100~\textsf{nm} was performed with a Langvin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
5~\textsf{ps}.  They server as starting point of the branching
trajectories. The branching trajectories were integrated by the
Leap-frog scheme without any temperature and pressure control.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electorstatic interaction in this
periodic system.  On each branching trajectory, to easy the analysis,
the conformation of the alanine dipeptide is projected on the
$\psi$--$\phi$ plane, as shown in Fig.~\ref{fig:tmp4}. The equilibrium
distribution of the conformation on the $\psi$--$\phi$ plane is given
in the Left plot of Fig.~\ref{fig:tmp4}.  It is clear the
conformations are populated in several clusters that are called
``metastable states'' in this research. For a clearer definition, we
manually devide the $\psi$--$\phi$ plane into 5 subregions. $A_1$ and
$A_2$ are corresponding to the alpha helix conformation. $B_1$ and
$B_2$ are corresponding
to beta sheet conformation. $C$ is corresponding to ill-folding.\\

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-begin-1.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-end-1.eps}
  \caption{The population of different metastable state. Left: the equilibrium state at $t=0$~\textsf{ps}. Right: The non-equilibrium stady state at $t=1200$~\textsf{ps}, i.e., fully relaxed when the electric field is switched on. The darker color indicates larger population on the $\psi$--$\phi$ phase space.}
  \label{fig:tmp4}
\end{figure}

For all testing cases, the maximum strength of the electric field
$E_{\textrm{max}}$ is chosen to be 1~\textsf{V/nm}.  Three values of
$t_{\textrm{init}}$ (10~\textsf{ps}, 100~\textsf{ps} and
500~\textsf{ps}) are chosen to study the response of the system to the
external electric field perturbation. The result is given in
Fig.~\ref{fig:tmp5}.  No matter how fast is the electric field turned
on, the beta sheet configuration, which presents at the beginning,
fades away as the system relaxes to the new equilibrium. The
population of alpha helix $A_1$ configuration grows from 45\% to 70\%.
alpha helix $A_2$  does not change a lot under the electric field.
The ill-folding configuration $C$ noticably grows from 0\% to 20\%.
\\

In the case of fastly switching on electric field
($t_{\textrm{init}}=10~\textsf{ps}$), the dynamics of the population
of the metastable sets shows three different time-scales. (1) From
$t=0$ to 10~\textsf{ps}, the population of $A_2$ fastly increases,
while the population of $A_1$ decreases with the same amount, so it is
reasonable to argue that there is a fast configurational change from
$A_1$ to $A_2$. Please see, in Fig.~\ref{fig:tmp5}, the green and read
peak arount $t = 10~\textsf{ps}$. At the same time, the population of
other metastable sets remains unchanged.  (2), From $t=10$ to
40~\textsf{ps}, the population of $B_2$ decreases by 0.15, and this
configuration almost disappers. Interestingly, the population of $A_1$
increases by the same amount, and the population of $A_2$ goes back to
the initial value. The populations of $B_1$ and $C$ start increasing
and decreasing, correspondingly. The speed of these changes are much
slower than that of $B_2$ and $A_1$.  (3) At the time scale of $t =
300~\textsf{ps}$, the population of $A_1$, $B_1$ and $C$ converge to
the non-equilibrium steady state (system with $E(\infty) = 1~\textsf{V/m}$).
This time is actually thirty times large than when the electric field
is fully switched on ($t_{\textrm{init}} = 10~\textsf{ps}$). This 
indicates that the internal time scale of $A_1$, $B_1$ and $C$
relaxation is around $300~\textsf{ps}$.\\

To further test the dynamics of metastable population, two simulations
of $t_{\textrm{init}} = 100~\textsf{ps}$ and $t_{\textrm{init}} =
500~\textsf{ps}$ are performed. The first time scale of $t =
10~\textsf{ps}$ disappers in both of the testing simulations. In the
simulation of $t_{\textrm{init}} = 100~\textsf{ps}$, the population of
$A$, $B_1$ and $C$ converges to the non-equilibrium steady state at
around $300~\textsf{ps}$. This further verifies internal time scale of
$A_1$, $B_1$ and $C$ presented in the testing case of
$t_{\textrm{init}} = 10~\textsf{ps}$.  The dynamics of $B_2$ is
faster: it decays to zero as fast as the external electric field is
switched on. In the case of $t_{\textrm{init}} = 500~\textsf{ps}$, the
growth of the external electric field is slower than the longest
internal time scale of system, which means that the changes of the
perturbation to the system is so slow that the system can catch up
with that. The numerical results show the same story: the relaxation
of all populations are around $500 \sim 600~\textsf{ps}$.
\\

@@

Let's consider an infinitely large NVE system (here we call it
"universe"). Some interesting process is happening in a small
subsystem. For equilibrium phenomena, this subsystem is the usually
considered as NVT (or NPT, $\mu$VT, depends on how the question is
raised) system.  In simulations, we only simulate the subsystem and
use the thermostat to mimic the embedment into the universe.  For
non-equilibrium simulation, we also want to simulate only the
subsystem of interest, but the physical meaning of directly applying a
thermostat designed for generating equilibrium ensemble is not clear
to a non-equilibrium process. Moreover, there is no unique way of
defining how the subsystem is coupled into the universe: It is subject
to how the perturbation comes into the subsystem, and how the energy
of the system dissipates into the universe. In the case of swithing on
electric field system, two limiting senarios are considered by us: (1)
the subsystem is isolated from the universe. That is to say the we
study the non-equilibrium behavior of a system, which conserves the
energy, volume and number of particle, under the turning on electric
field.

(1), the
electric field is applied to the whole universe, the correctness of
considering a subsystem is tested by varying the size of the
subsystem, and investigating the convergence of the result.

(2), the
perturbation is applied locally, and the extra energy dissipates into
the universe infinitely fast. That is to say the subsystem is coupled
to a thermostat and barostat with infinitely high
efficiency. Practically, it is impossible to simulation with a
``infinitely fast'' thermostat and barostat. We locally truncate the
system with radius $R_{\textrm{ex}} = 1~\textsf{nm}$ (see
Fig.~\ref{fig:tmp2} for explanation), and couple it to very sufficient
thermo- and barostat. We use a Langvin thermostat coupled to the
thermostating region with a time-scale of $\tau_{T} =
0.1~\textsf{ps}$, and a Parrinello-Rahman barostat with a time-scale
of $\tau_{P} = 2.0~\textsf{ps}$.

The non-equilibrium steady temperature and pressure of the
non-thermostated simulation are $330$ K and -500 Bar,
correspondingly. Those of the thermostated sytem are $300$ K and 1
Bar, correspondingly. Therefore, we can see from Fig.~\ref{fig:tmp6}
that the relaxation of the thermostated system is faster than the
non-thermostated system. The three time-scale of metastable population
dynamics in the thermostated system is consistent with that of the
non-thermostated system. This means that thermostating, in this case,
does not qulitativly change the population dynamics of metastable
states.

>>>>

\section{Alanine dipeptide
  under turning on electric field}
\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig.ala/ele.field/field.eps}
  \caption{The magnitude of the external electric field as
    a function of time.}
  \label{fig:tmp3}
\end{figure}

We study the response of the configuration of the alanine dipeptide
with respect to a turning on electric field. The direction of this
field is along the $x$ direction and the magnitude with respect to time
is plotted in Fig.~\ref{fig:tmp3}. From $t=0$ to
$t=t_{\textrm{init}}$, the strength of the field grows linearly, while
after $t=t_{\textrm{init}}$, the strength is kept constant at
$E_{\textrm{max}}$.  The size of the system is $2.67\times 2.67\times
2.67\, \textsf{nm}^3$, with one alanine dipeptide solved in 644 TIP3P
water molecules. Firstly, an equilibrium NVT simulation of
100~\textsf{nm} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$) coupled to the system for constant
temperature.  Along the trajectory, configurations were taken every
5~\textsf{ps}.  They server as starting point of the branching
trajectories. The branching trajectories were integrated by the
Leap-frog scheme without any temperature and pressure control.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by Gromacs.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system.  On each branching trajectory, to easy the analysis,
the conformation of the alanine dipeptide is projected on the
$\psi$--$\phi$ plane, as shown in Fig.~\ref{fig:tmp4}. The equilibrium
distribution of the conformation on the $\psi$--$\phi$ plane is given
in the Left plot of Fig.~\ref{fig:tmp4}.  It is clear the
conformations are populated in several clusters that are called
``metastable states'' in this research. For a clearer definition, we
manually divide the $\psi$--$\phi$ plane into 5 subregions. $A_1$ and
$A_2$ are corresponding to the alpha helix conformation. $B_1$ and
$B_2$ are corresponding
to beta sheet conformation. $C$ is corresponding to ill-folding.\\

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-begin-1.eps}
  \includegraphics[width=0.45\textwidth]{fig.ala/ext.mode1.100.Ex.01.00.t1200ps/fig-end-1.eps}
  \caption{The population of different metastable state. Left: the equilibrium state at $t=0$~\textsf{ps}. Right: The non-equilibrium study state at $t=1200$~\textsf{ps}, i.e., fully relaxed when the electric field is switched on. The darker color indicates larger population on the $\psi$--$\phi$ phase space.}
  \label{fig:tmp4}
\end{figure}

For all testing cases, the maximum strength of the electric field
$E_{\textrm{max}}$ is chosen to be 1~\textsf{V/nm}.  Three values of
$t_{\textrm{init}}$ (10~\textsf{ps}, 100~\textsf{ps} and
500~\textsf{ps}) are chosen to study the response of the system to the
external electric field perturbation. The result is given in
Fig.~\ref{fig:tmp5}.  No matter how fast is the electric field turned
on, the beta sheet configuration, which presents at the beginning,
fades away as the system relaxes to the new equilibrium. The
population of alpha helix $A_1$ configuration grows from 45\% to 70\%.
alpha helix $A_2$  does not change a lot under the electric field.
The ill-folding configuration $C$ noticeably grows from 0\% to 20\%.
\\

In the case of fastly switching on electric field
($t_{\textrm{init}}=10~\textsf{ps}$), the dynamics of the population
of the metastable sets shows three different time-scales. (1) From
$t=0$ to 10~\textsf{ps}, the population of $A_2$ fastly increases,
while the population of $A_1$ decreases with the same amount, so it is
reasonable to argue that there is a fast configurational change from
$A_1$ to $A_2$. Please see, in Fig.~\ref{fig:tmp5}, the green and read
peak around $t = 10~\textsf{ps}$. At the same time, the population of
other metastable sets remains unchanged.  (2), From $t=10$ to
40~\textsf{ps}, the population of $B_2$ decreases by 0.15, and this
configuration almost disappears. Interestingly, the population of $A_1$
increases by the same amount, and the population of $A_2$ goes back to
the initial value. The populations of $B_1$ and $C$ start increasing
and decreasing, correspondingly. The speed of these changes are much
slower than that of $B_2$ and $A_1$.  (3) At the time scale of $t =
300~\textsf{ps}$, the population of $A_1$, $B_1$ and $C$ converge to
the non-equilibrium steady state (system with $E(\infty) = 1~\textsf{V/m}$).
This time is actually thirty times large than when the electric field
is fully switched on ($t_{\textrm{init}} = 10~\textsf{ps}$). This 
indicates that the internal time scale of $A_1$, $B_1$ and $C$
relaxation is around $300~\textsf{ps}$.\\

To further test the dynamics of metastable population, two simulations
of $t_{\textrm{init}} = 100~\textsf{ps}$ and $t_{\textrm{init}} =
500~\textsf{ps}$ are performed. The first time scale of $t =
10~\textsf{ps}$ disappears in both of the testing simulations. In the
simulation of $t_{\textrm{init}} = 100~\textsf{ps}$, the population of
$A$, $B_1$ and $C$ converges to the non-equilibrium steady state at
around $300~\textsf{ps}$. This further verifies internal time scale of
$A_1$, $B_1$ and $C$ presented in the testing case of
$t_{\textrm{init}} = 10~\textsf{ps}$.  The dynamics of $B_2$ is
faster: it decays to zero as fast as the external electric field is
switched on. In the case of $t_{\textrm{init}} = 500~\textsf{ps}$, the
growth of the external electric field is slower than the longest
internal time scale of system, which means that the changes of the
perturbation to the system is so slow that the system can catch up
with that. The numerical results show the same story: the relaxation
of all populations are around $500 \sim 600~\textsf{ps}$.
\\

@@

Let's consider an infinitely large NVE system (here we call it
"universe"). Some interesting process is happening in a small
subsystem. For equilibrium phenomena, this subsystem is the usually
considered as NVT (or NPT, $\mu$VT, depends on how the question is
raised) system.  In simulations, we only simulate the subsystem and
use the thermostat to mimic the embedment into the universe.  For
non-equilibrium simulation, we also want to simulate only the
subsystem of interest, but the physical meaning of directly applying a
thermostat designed for generating equilibrium ensemble is not clear
to a non-equilibrium process. Moreover, there is no unique way of
defining how the subsystem is coupled into the universe: It is subject
to how the perturbation comes into the subsystem, and how the energy
of the system dissipates into the universe. In the case of switching on
electric field system, two limiting scenarios are considered by us: (1)
The subsystem is isolated from the universe: The electric field is
applied to a system that conserves the energy, volume and number of
particle (i.e. a NVE subsystem in short).  The isolation here means
that the subsystem does not exchange energy, volume and number of
particle with the universe, but the perturbation (electric field) does
come in to the subsystem.  The periodic boundary condition is applied
here only to avoid the difficulty of defining a experimental
meaningful boundary in simulation.  In this scenario, the size the
subsystem is somewhat arbitrary (it depends on experimental setting),
therefore, it is important to consider the convergence of the
simulation result with respect to the size of the subregion.  (2) The
system exchange the energy, volume and particle with the universe, and
the universe is so big that the perturbation in the subsystem will not
kick the universe from equilibrium.  In practice, it is impossible to
simulate the infinite large universe, Instead, we want to simulate the
considerably smaller subsystem instead.  The idea is to divide the
simulation region into two parts (see Fig.~\ref{fig:tmp2}): the first
is called \emph{Newtonian dynamics region} that is a region around the
molecule of interest, where the dynamics is Newtonian. Note that the
molecular trajectories in the Newtonian dynamics region are the same
as if it were couple to the universe, only when a proper boundary
condition is provided. Strictly speaking, the boundary should be
provided by the out side infinitely large universe, which immediately
leads back to the numerical difficulty of simulating a infinite large
system. However, we notice that the detailed information
(i.e. position and velocity) of the every molecule outside the
Newtonian dynamics region is not relevant to the non-equilibrium
behavior of the molecule of interest, because we are only interest in
the non-equilibrium averages, which does not depends on the individual
trajectory, but on the trajectorial ensemble sampled by the
non-equilibrium simulation.  Therefore, the universe, as a boundary to
the Newtonian dynamics region, can be replace by
an efficiently thermo- and barostated subsystem, as indicated by the gray
part of Fig.~\ref{fig:tmp2}. 
% That is to say the subsystem is coupled to a
% thermostat, a barostat and a particle bath with infinitely high
% efficiency. Practically, it is impossible to simulation with a
% ``infinitely fast'' thermostat, barostat and particle bath.
We indicate the radius of the Newtonian dynamics region by
$R_{\textrm{ex}} = 1~\textsf{nm}$ (see Fig.~\ref{fig:tmp2} for
explanation), and couple it to very sufficient thermo- and
barostat. We use a Langevin thermostat coupled to the thermostating
region with a time-scale of $\tau_{T} = 0.1~\textsf{ps}$, and a
Parrinello-Rahman barostat with a time-scale of $\tau_{P} =
2.0~\textsf{ps}$.

The non-equilibrium steady temperature and pressure of the
non-thermostated simulation are $330$ K and $-500$ Bar,
correspondingly. Those of the thermostated system are $300$ K and 1
Bar, correspondingly. Therefore, we can see from Fig.~\ref{fig:tmp6}
that the relaxation of the thermostated system is faster than the
non-thermostated system. The three time-scale of metastable population
dynamics in the thermostated system is consistent with that of the
non-thermostated system. This means that thermostating, in this case,
does not qualitatively change the population dynamics of metastable
states.
